\chapter{Intelligent tutoring system}
\label{ch:its-implementation}
\glsreset{its}
%(CONTEXT): background for less specialized readers and establish or recalls the importance of the problem
An important aspect of education in the paved path methodology is its efficiency.
Educational activities should not keep the developer from their responsibilities for longer than necessary.
% (NEED): motivates the audience by stating the difference between the desired and actual situation
% ==>shortcomings and goal from my perspective
The efficiency of training on the \gls{scw} platform is lacking.
Because every individual is presented with the same exercises, they often receive training that is too repetitive or too challenging.
%(The TASK) what did I do?
%=> IRT, ITS
I designed an \gls{its} that recommends exercises to each individual at any point in time to provide them with a more appropriate learning pace.
%When evaluating online learning activities, the focus is often wrongfully on completion and dropout rates~\cite{hadi2016driving}. 
%It is perfectly possible, likely even, that learning took place despite the user not completing the course.
%The goal of the \gls{its} should hence be to maximize the opportunity for learning to take place.
% OBJECT
In this chapter I present the design of the \gls{its} and discuss the used techniques in its implementation in more detail.
%In this chapter we start with an overview of the design of the \gls{its} in Section~\ref{sec:design}.
%In the rest of the chapter we explain the used techniques in more detail.
%First, the collaborative filtering algorithm is detailed in Section~\ref{sec:collab}.
%Next, the psychometric models are discussed in Section~\ref{sec:calib}.

\summarybox{
I designed an \gls{its} that consists of three main algorithms, one for exercise selection, and one each for estimation of user ability and exercise difficulty.
Exercise selection is achieved through a \gls{cf} algorithm adapted to learning systems.
In such an algorithm a target user's preference for an exercise is predicted from the preferences of like-minded users.
Through the use of the psychometric model of \gls{irt} both estimation of user ability and exercise difficulty can be done at once.
A sanitized data set of over 9 million solved exercises is used to calibrate these algorithms making up the \gls{its}.
}

% SECTIONS
\input{03-education/02-intelligent-tutor/sections/01-design}
\input{03-education/02-intelligent-tutor/sections/02-collaborative-filtering}
\input{03-education/02-intelligent-tutor/sections/03-calibration}
\input{03-education/02-intelligent-tutor/sections/04-data}