\chapter{Discussion and perspectives}
\label{ch:its-evaluation}

In the previous chapter, I described the goal and set-up of each experiment, and reported their findings.
In this chapter, I summarize the findings and explain the learned lessons.
I describe how these results can be used to provide better support and training to developers. 

\summarybox{
The \gls{2pl} model has shown that some infamous vulnerability types that are typically considered high priority are relatively easy to find and fix in training.
This is evidence of a gap between knowledge and practice.
The paved path methodology, together with more usable developer tools, could help developers apply their knowledge better.

For use in the \gls{its}, the \gls{knn} baseline algorithm is the most accurate \gls{cf} algorithm.
Baselines have proven to be accurate for the data of the platform, most likely because many users show consistent bias in their ratings.
This is related to the current item selection, that is often consistently right or consistently wrong, based on the ability level of the user.
In literature, model-based algorithms are usually more accurate, but their advantages of dealing with data sparsity, scalability, synonyms, and implicit data are not applicable to the data set of the \gls{scw} training platform.

The results of the \gls{2pl} model and the \gls{its} will be gradually implemented in the training platform.
In the future the \gls{its} will be improved to use data gathered from other developer tools, as described in a published patent by \gls{scw}.
}

\input{03-education/04-evaluation/sections/04-discussion}
\input{03-education/04-evaluation/sections/05-perspectives}