\chapter{Experiments}
\label{ch:its-experiments}

Several techniques and algorithms have been combined in the design of the \gls{its}.
With the implementation and evaluation of each of these, new insights were gained into the mental model of the developer. 
In this section, I describe the four experiments I conducted to evaluate and fine-tune each component in the \gls{its}.

\summarybox{
In the first experiment, I evaluated the results of the Rasch model through statistical analyses.
I found that the assigned difficulty of an exercise on the \gls{scw} platform does not accurately represent the actual difficulty as experienced by the users.
The statistical tests show that the actual difficulty does depend on the framework of the exercise, the category of the vulnerability, and the presentation of the exercise to the user. 

The second experiment showed that out of the tested approximation methods, the one invented in this work is the best to achieve a fast ability estimates at each point in time.
The mean error stays below 10\% even after more than 200 completed challenges.

In the final two experiments, I tested the performance of different \gls{cf} algorithms and their adaptations to learning systems.
The adaptations caused an increase in predication accuracy between 3.9\% and 13.7\%, depending on the algorithm.
The best performing algorithm in the end is the \gls{knn} baseline algorithm, achieving a mean absolute error of 0.4206.
}

\input{03-education/03-experiments/sections/00-rasch}
\input{03-education/03-experiments/sections/00-stepsize}
\input{03-education/03-experiments/sections/01-cf}
\input{03-education/03-experiments/sections/02-learning}

