\chapter{Sensei}
\label{ch:sensei}

% Context
The development of the Sensei \gls{ide} plugin started in 2016, when dr. Matias Madou and Nathan Desmet founded the company Sensei Security.
I joined this company, that later would merge with \gls{scw}, as an intern a few months later.
% need and task in one
When I started my research in 2017, I set forth to discover how this tool could be used most effectively, to evaluate its concepts and features, and to help direct its design.
In this chapter, I describe the Sensei \gls{ide} plugin and discuss the lessons we learned during the implementation and testing of the tool.

\summarybox{
The first iteration of the Sensei rule editor was a \gls{gui} containing many input fields to allow fast customization of rules.
It was used by us to create hundreds of rules for customers and developer communities which frequently resulted in the need for extra features.
Some of these features are useful for improving the context awareness of Sensei and its usability, other features fell out of use.
Eventually, through the addition of these many features, the rule editor became too cluttered and unclear.

As a more flexible alternative, Nathan Desmet, principal engineer at \gls{scw}, and I designed a new formatting language based on \gls{yaml} syntax that allows rule writers to quickly and effectively create rules and quick-fixes.
The rules include several features to improve their usability and add support for libraries and for design flaws.
}

\input{04-tools/03-sensei/sections/00-installation}
\input{04-tools/03-sensei/sections/01-rules}
\input{04-tools/03-sensei/sections/02-rule-features}
%\input{04-tools/03-sensei/sections/02-install}
%\input{04-tools/03-sensei/sections/02-plugin}

