\part{Education}
\label{p:education}
\epigraph{Pass on what you have learned. Strength. Mastery. But weakness, folly, failure also. Yes, failure most of all. The greatest teacher, failure is.}{\textit{Yoda}}

\clearpage
\section*{Introduction}
%(CONTEXT): background for less specialized readers and establish or recalls the importance of the problem
Education in the paved path methodology should be deliberate and keep the developer experience in mind.
The \gls{scw} training platform is a good resource for education in the paved path methodology. 
It provides online training through defensive secure coding exercises in many different programming languages and frameworks and its gamification and interactivity make it fun and usable.
%
%% (NEED): motivates the audience by stating the difference between the desired and actual situation
%% ==>shortcomings and goal from my perspective
Despite the focus on developers and its many usability features, there is still a significant part of the user base that only follows a minimal amount of training.
Users follow one of the predetermined courses, and it is likely that the pacing of these courses does not fit their needs.
Users get bored due to too much repetition, or frustrated because the content is moving too fast.

%(The TASK) what did I do?
%=> IRT, ITS
I created an \gls{its} by combining psychometric models with techniques from computer science to recommend exercises to each individual at any point in time. 
A customized recommendation like this is more likely to keep the developer engaged and allows meaningful learning to take place.

I start Part~\ref{p:education} by giving an overview of the \gls{scw} training platform, its features, and the different types of exercises in Chapter~\ref{ch:scw}.
I then describe the design of the \gls{its} and explain the used techniques in more detail in Chapter~\ref{ch:its-implementation}.
The last chapter in this part provides an evaluation of the \gls{its} before offering some perspectives that remain future work.

%%(FINDINGS) what did I find?
%\todo[inline]{what did I find?}
%(TODO) The recommendation system was able to predict the learning outcome of challenges in historical data with XX\% accuracy. 
%
%%(CONCLUSION) what does this mean for the need?
%\todo[inline]{what does this mean for the need? --> will there be better engagement? More learning?}
%(TODO) Recommending only challenges with a high probability of a positive learning outcome greatly improves the efficiency of the education.
%It also reduces the mismatch in learning pace and keeps developer engaged longer.
%The \gls{its} also allows us to identify defect exercises, as well as promote new content.
%
%%(PERSPECTIVES) what now?
%%=> integrations
%%=> ecosystem
%There are still opportunities to improve the \gls{its} in the future. 
%The psychometric models allow us to take into account the ability of a user, however, we expect the agility of a user to be important as well.
%It is also possible that recommendations can be improved by taking into account data from other developer tools, such as the issue tracking system, the code repository, and security tools. 
%The current implementation only uses data from the training platform itself.

\input{03-education/00-scw/00-scw}
\input{03-education/02-intelligent-tutor/02-intelligent-tutor}
\input{03-education/03-experiments/03-experiment}
\input{03-education/04-evaluation/04-evaluation}
