\part{Tools}
\label{p:tools}
\epigraph{Do, or do not. There is no try.}{\textit{Yoda}}

\clearpage
\section*{Introduction}
% (CONTEXT): background for less specialized readers and establish or recalls the importance of the problem
% ==> state and vision without me

Despite growing efforts to educate developers, they still frequently make mistakes in practice.
Because security experts are understaffed and unable to assist each developer, security tools have become part of the developer's arsenal.

% (NEED): motivates the audience by stating the difference between the desired and actual situation
% ==>shortcomings and goal from my perspective
However, security tools are not designed with the developer in mind and are often a big inhibitor of their productivity. As a result developers dislike and often disable these security tools. 


% (The TASK) What did I do?
We designed and implemented an \gls{ide} plugin called Sensei in line with the paved path methodology.
Sensei has a heavy focus on developer usability and productivity.
It checks compliance of code being written to a set of coding guidelines, similar to an as-you-type spell checker.
Upon detected violations, it offers remediation guidance and additional information to the developer.
Sensei allows developers and security experts to develop customized rule sets to be enforced, specific to each project.

% (OBJECT)
In this part I first describe the evolution of traditional security tools in Chapter~\ref{ch:goals}.
I explain where they are lacking as a tool to support development in the paved path methodology and what goals and requirements to set instead.
In Chapter~\ref{ch:sensei}, I describe the Sensei \gls{ide} plugin, its design, and its features.
I continue by describing the experiments and observations of the tool in Chapter~\ref{ch:experiments}.
In Chapter~\ref{ch:tool-eval}, I discuss the findings of the experiments and the lessons learned and also offer some perspectives that remain future work.

% (FINDINGS)
%We observed that developers stay engaged with Sensei and heavily use its remediation guidance.
%Developers also write their own rules for various purposes other than security.
%
%% (CONCLUSION)
%Sensei is able to overcome some of the inhibitors of developer productivity caused by traditional security tools such as slow scanning speeds, false positives, and lack of remediation guidance.
%
%% (PERSPECTIVES)
%Although this tool is already a big step in the right direction, developer usability can still be improved by making it more user aware, and by improving its usability on legacy code bases.
%Another challenge that remains future work is the automated creation of custom guidelines.

\input{04-tools/00-goals/00-goals}
\input{04-tools/03-sensei/03-sensei}
\input{04-tools/04-experiments/04-experiments}
\input{04-tools/05-evaluation/05-evaluation}

\part{Related work}
\label{p:related}
\epigraph{Read them have you? Page-turners they were not.}{\textit{Yoda}}

\input{04-tools/07-related/07-related}