\chapter{Goals and requirements}
\label{ch:goals}

% CONTEXT
Traditional security tools are an important part of the \gls{sdlc}, and will remain so in the foreseeable future.
However, they hinder productivity and do not integrate well in the workflows of developers.
To provide developers with more suitable tools, it is critical to understand the goals and the shortcomings of traditional security tools.
% TASK
In this chapter, I briefly describe how security practices and tools evolved over time.
I explain the disconnect with developer workflows and how the paved path methodology can improve this situation.

\summarybox{
% Findings
The goal of traditional security tools is automation of security testing.
In order to keep up with the ever increasing speed of software development, they are \textit{shifting left} in the \gls{sdlc}, towards the development phase.
As a result they are being integrated in developer tools.
However, they are still fundamentally using a reactive approach, scanning (partly) completed code and its calling context for \glspl{vulnerability}.
To fix detected \glspl{vulnerability}, developers often have to go back to the code (potentially long) after it was initially developed.
With tools in the paved path methodology the goal is a preventative approach.
Guidelines are enforced regardless of context as the code is being written.
% Conclusion
This helps the developer write secure code from the start, improving their productivity.
As a result, code fragments are being secured, even if no current calling context exists that leads to an \gls{exploit}, they are being secured for future use.
}

\input{04-tools/00-goals/sections/01-shift-left}