\thispagestyle{empty}

\vspace*{\fill}
\begin{center}
\begin{minipage}{.6\textwidth}
\emph{Always in motion the future is -- Yoda}
\end{minipage}
\end{center}
\vfill % equivalent to \vspace{\fill}

%\vfill\vfill

\cleardoublepage

\chapter*{Acknowledgements}
\label{ch:ack}

We acquire many new skills during life, and they are taught to us by many different people.
Ever since I became a parent, I have realised just how much impact some of these people have had on me.
Being a dad comes rather natural to me, and I can only attribute this to the many great examples I have had in my life of loving parents and grandparents.
In the same way, I want to attribute the academic success of this work to the many great examples I have encountered during these last few years.

\subsubsection*{My thanks to...}

\paragraph{Pieter Vandenbossche}
for standing on top of the desks, reading Latin texts in funny voices, and conducting an imaginary orchestra while playing Carmina Burana in class.
My Latin teacher in high school taught me that interest and passion for a subject are infectious and to look for those teachers and students that can infect me.

\paragraph{Bjorn De Sutter}
for being such an infectious teacher.
I found that paying attention to his classes took no effort at all and so I happily attended all possible elective courses taught by him.
It is because of Bjorn that I found my way into the world of security.

\paragraph{Matias Madou}
for teaching me about hard work and determination, but also about not always taking myself too seriously.
It was based on his advice that I applied for a grant and started this research.
Matias is not only the world's okayest boss, but also a great mentor.

\paragraph{my examination board}
for their great feedback on both my research and this dissertation.
I appreciate all the time and effort they spent evaluating this work.

\paragraph{the technical advisory board at SCW}
for their feedback on the intelligent tutor both in meetings and over mail.
Brian's continued interest and feedback have been a source of confidence, and Chenxi's advice to \textit{do the dumb thing first} will be a lesson that sticks with me for a long time.

\paragraph{the students, teaching staff, and security professionals}
who participated in the experiments and interviews in this dissertation.

\paragraph{my (former) colleagues at SCW and SysLab}
who made coming to work fun and relaxing.
I love the atmosphere at the Bruges office, and the culture at SCW in general.
Mondays would not be as fun without breakfast at work.
And, it is always a pleasure to travel to Gent and catch up with my coworkers at the iGent tower.
I already look forward to the next game night.

\paragraph{Stéphanie}
for helping me improve my graphs, and for all her support and friendship.

\paragraph{Brysen, Downey, and Tim}
for their feedback, insights, and assistance during my work, and for the countless evenings playing games to help me unwind.

\paragraph{Laurens and Wouter}
for the many hours studying together, working on group projects, and of course the many hours of playing board games.

\paragraph{Gillis}
for his friendship, and support, and for always being there when I need a sympathetic ear.
I wish everyone a friend as kind, down-to-earth, and awesome as Gillis.

\paragraph{my family}
for the unconditional support and love through all these years and for giving me a warm place to call home.
When I am in need, there is never any doubt that my parents, my sister Laura, and her husband Bram are all ready to help however they can.

\paragraph{Sarah}
for being a firm believer, and a never ending source of support through the sometimes difficult times.
Sarah and our daughter Suzanne are my biggest fans.
I could not have finished this without them.