\chapter{Related work}
\label{ch:related}

\epigraph{Read them have you? Page-turners they were not.}{\textit{Yoda}}

Software security is a relatively new field~\cite{mcgraw2004software}, but many tools and practices have already been developed that have caused great advancements.
Many guides exist to help you decide which tools are appropriate for your project.
In this chapter, I want to extend these guides so that you are better equipped to estimate the strengths of a tool for use in the paved path methodology.
I describe commonalities between tools in different phases of the \gls{sdlc} and how they can be deployed effectively.

For future reference, Appendix~\ref{app:battlecards} contains a number of battlecards with my thoughts on a few more tools.

\summarybox{
Security begins even before code is written.
Laws, legislations, and consumer demands all impact how much attention is given to security.
Besides lightweight linter tools, developers can also find help to produce secure code from patterns, libraries, and frameworks.
In the build phase, the use of new methodologies has driven the automation of building executables and installing dependencies, which has made it easier to test for use of vulnerable components.

Most security practices, however, take place in the test phase.
Many code review practices and tools exist, most of which allow customization of the rules through one of three methods: an \gls{api}, a custom query language, or a formatting language such as \gls{xml} or \gls{yaml}.
Finally, in the release phase, recent advancements in \gls{iac} have made it easier to securely deploy applications and manage infrastructure.
}

\input{05-part-three/07-related/sections/00-plan}
\input{05-part-three/07-related/sections/01-develop}
\input{05-part-three/07-related/sections/02-build}
\input{05-part-three/07-related/sections/03-test}
\input{05-part-three/07-related/sections/04-release}