\chapter{Experiments and observations}
\label{ch:experiments}

Based on our own experience and anecdotal evidence, we have set goals and priorities during the development of Sensei.
During my research, I conducted various experiments and interviews to asses these priorities, and to evaluate the various features as described in the previous chapter.
In this chapter I describe the goal of each experiment, its set-up, and report the findings.

\summarybox{
I conducted a controlled user experiment that showed Sensei markings are easy to understand and quick-fixes are applied frequently.
On average, subjects that were using Sensei to develop security critical features, had an increase in development time of about 11\%.

Interviews with security professionals showed that the tool can be used effectively in a professional setting to detect and remediate security problems.
They describe the customization of recipes as being easier and faster compared to other security tools.
Nonetheless, usability experiments showed that the \gls{yaml} code used for recipes can be overwhelming and that all users prefer the \gls{ui} view of the new recipe editor.
Creating new recipes was more successful if the users followed along with documentation, or if they looked at example recipes first, creating recipes from scratch can still be difficult.

The biggest disadvantages of Sensei compared to other tools are its lack of reported metrics and its poor integration in the \gls{cicd} pipeline.
}

%\todo[inline]{compare results to those of rasch model}

\input{04-tools/04-experiments/sections/00-experiment}
\input{04-tools/04-experiments/sections/01-user-testing}
\input{04-tools/04-experiments/sections/02-trial}
\input{04-tools/04-experiments/sections/03-trial}
%\input{04-tools/04-experiments/sections/04-marketplace}

